\chapter{Overview}
\section{\acl{pam}}
The signal is given by ($g(t)$ is basic pulse shape, $a \in \mathcal{A}$ is a signal
point):
\begin{equation}
    s(t) = \sum_{k=-\infty}^\infty a[k] g(t-kT)
\end{equation}

\subsection{\acl{ask}}
Signal points on the real axis either only positive (unary) or symmetrical (binary).

\subsection{\acl{psk}}
Signal points on the unit circle.

\subsection{\acl{qam}}
Signal points spaced regularly on the complex plane.

\subsection{\acl{oqam}}
\ac{qam} with the quadrature branch delayed by $T/2$ thus only the in-phase or the
quadrature part changes. This yields a signal without zero crossing, which yields a
more constant envelope and thus a smaller crest factor.

\subsection{\acl{msk}}
\ac{oqam} with $M=4$ and a truncated $\cos$ (i.e. only the positive part centered around
zero) as $g$. Yields a constant envelope and thus $\zeta=1$.

\subsubsection{Pros}
Constant envelope \ac{qam}

\subsubsection{Cons}
Half the spectral efficiency of \ac{qam}. Infinite Bandwidth.

\subsection{\acl{gmsk}}
Filtering of \ac{msk} with a gaussian filter to achieve a smaller bandwidth

\subsubsection{Cons}
Nonlinear.

\subsection{\acl{dpsk}}
Information not in absolute phase but relative to last phase.

\ac{dcpsk} receiver: \ac{psk} receiver followed by differentiation.

\ac{dpsk} receiver: first differentiation then a \ac{psk} demodulator.

\subsubsection{Pros}
Can handle non coherent phase.

\subsubsection{Cons}
Worse performance than coherent schemes, binary \ac{dpsk} has equal performance.

\section{\acl{fsk}}
Information is encoded in the instantaneous frequency. We can add a additional phase offset
to achieve continuous \ac{fsk} in contrast to discontinuous \ac{fsk}.

\section{Dispersive Channels}
\subsection{\acl{le} / \acl{zfle}}
The \ac{isi} is force to zero at the sampling time steps.

\subsubsection{Pros}
No \ac{isi}.

\subsubsection{Cons}
Non white noise.


\subsection{\acl{dfe}}
The last symbols are used to subtract the additive noise (smearing) caused by the channel.

\subsubsection{Pros}
Simple implementation, better performance than \ac{le}.

\subsubsection{Cons}
Coding can not be considered. Error will propagate over time.

\section{\acl{ofdm}}
The bandwidth is split equally into $D$ subchannels. The data symbols are interpreted to be
frequency domain symbols. Loading optimizes the datarate per subchannel to achieve better
performance.
